### 2018-02-01 includes/appendix.tex

We will make the folowing simplifications for our proofs:
\begin{itemize}
  \item
    Since many of the operators can be expressed in terms of other, our
    proofs will only consider a subset of the operators. In particular we will
    drop the following:
    \begin{itemize}
      \item \texttt{||} (boolean or)
      \item \texttt{=} for boolean expressions
      \item \texttt{<=}, \texttt{>}, \texttt{>=}
      \item \texttt{str.prefixof}, \texttt{str.suffixof}, \texttt{str.at}
      \item \texttt{str.contains}
      \item \texttt{re.range}, \texttt{re.+}, \texttt{re.allchar}, \texttt{re.all}
    \end{itemize}

  \item
    In the above, we assume that \texttt{str.at} is expressible using
    \texttt{substring} although the semantics slightly different when the index
    is out of bounds.

  \item
    We will assume that our program instances are a single assertion. Multiple
    assertions can be expressed by combining them using \texttt{and}.
\end{itemize}

@@

The proofs all follow a very similar structure. For each transformation
\textit{Trans}, we show that a problem instance $P$ is satisfiable by a model $m$
if and only if $\textit{Trans}(P)$ is satisfiable by $\textit{Trans}(m)$, where
$\textit{Trans}(M)$ is the transformation applied to all the constants in the
model $m$.

@@

For transformations that have an inverse, such as \textit{Translate}
and \textit{Reverse} this also shows that if the transformed problem is
satisfiable, then the original problem is satisfiable. This proves that these
transformations are equisat transformations. \textit{Multiply} does not have an
inverse and hence it only guarantees to take satisfiable problems to satisfiable
problems. This can be seen in the problem instance $y<x<y+1$ which is UNSAT, but
multiplying by 2 transforms the problem into $y<x<y+2$ which is SAT.

\subsection{Translate}
We now prove that translate transforms programs into equisatisfiable
programs. From above, we only need to show that for every expression $e$,
$\textit{Translate}(\eval (e)) = \eval (\textit{Translate}(e))$. For now, we
assume that \textit{Translate} fixes the digit characters pointwise and we prove
the stronger fact that $\textit{Translate}$ also fixes the digits pointwise in
any string expression.

@@

We end our discussion on \textit{Reverse} by nothing that although it disallowed
\texttt{str.replace} only because it didn't have a natural notion of a reverse
in the theory we were working with. If we expand our theory to also include an
operator that replaces the last instance, we can transform problems that include
\texttt{str.replace}.

>>>>

To shorten our proofs, we will make the following simplifications:
\begin{enumerate}
  \item
    Since many of the operators can be expressed in terms of other, our
    proofs will only consider a subset of the operators. In particular we will
    drop the following:
    \begin{itemize}
      \item \texttt{||} (boolean or)
      \item \texttt{=} for boolean expressions
      \item \texttt{<=}, \texttt{>}, \texttt{>=}
      \item \texttt{str.prefixof}, \texttt{str.suffixof}, \texttt{str.at}, \texttt{str.contains}
      \item \texttt{re.range}, \texttt{re.+}, \texttt{re.allchar}, \texttt{re.all}
    \end{itemize}

  \item
    In the above, we assume that \texttt{str.at} is expressible using
    \texttt{substring} although the semantics slightly differ when the index is
    out of bounds.

  \item
    We will assume that our program instances are a single assertion. Multiple
    assertions can be expressed by combining them using \texttt{and}.
\end{enumerate}

@@

The proofs all follow a very similar structure. For each transformation
\textit{Trans}, we show that a problem instance $P$ is satisfiable by a model $m$
if and only if $\textit{Trans}(P)$ is satisfiable by $\textit{Trans}(m)$, where
$\textit{Trans}(m)$ is the transformation applied to all the constants in the
model $m$.

@@

For transformations that have an inverse, such as \textit{Translate} and
\textit{Reverse} this also shows that if the transformed problem is satisfiable,
then the original problem is satisfiable, proving that they are equisat
transformations. \textit{Multiply} does not have an inverse and hence it only
guarantees to take satisfiable problems to satisfiable problems. This can be
seen in the problem instance $y<x<y+1$ which is UNSAT, but multiplying by 2
transforms the problem into $y<x<y+2$ which is SAT.

\subsection{Translate}
We now prove that \textit{Translate} transforms programs into equisatisfiable
programs. From above, we only need to show that for every expression $e$,
$\textit{Translate}(\eval (e)) = \eval (\textit{Translate}(e))$. For now, we
assume that \textit{Translate} fixes the digit characters pointwise and we prove
the stronger fact that $\textit{Translate}$ also fixes the digits pointwise in
any string expression.

@@

We end our discussion on \textit{Reverse} by nothing that although it disallowed
\texttt{str.replace}, it only did so because it didn't have a natural notion of
a reverse in the theory we were working with. If we expand our theory to also
include an operator that replaces the last instance, we can transform problems
that use \texttt{str.replace}.
